WASP-4 is Accelerating Towards the Earth

### 2020-01-29 paper/ms.tex
update paper
ready for JNW!

\begin{abstract}
  Space and ground-based transit observations of the hot Jupiter
  WASP-4b have recently shown that its orbital period is decreasing.
  The period change could be caused by tidal orbital decay, orbital
  precession, or light-travel time effects.  We present new radial
  velocity measurements of WASP-4, acquired with Keck-HIRES.  The data
  show that the system is accelerating towards the Earth at
  $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm m}\,{\rm
  s}^{-1}\,{\rm yr}^{-1}$.  The implied period decrease probably
  explains WASP-4's changing orbital period as a light-travel time
  effect.  Combining the radial velocities with new limits from
  speckle imaging, we find that the system's acceleration is likely
  caused by a 10-200$\,M_{\rm Jup}$ companion with semi-major axis
  between 10-100$\,$AU.  The statistics from
  \citet{knutson_friends_2014} imply that about 1 in 6 hot Jupiters
  are expected to show similar period decreases as WASP-4, due to
  acceleration by outer companions.  We find that these period
  decreases will be increasingly measureable in coming years, because
  the precision of period derivative measurements scales as the fourth
  power of the observing baseline.  Continued radial velocity
  monitoring of hot Jupiters is therefore essential to distinguish
  tidal orbital decay from line-of-sight accelerations.
\end{abstract}

@@

Indirect studies of age and angular momentum indicators including the
hot Jupiter semi-major axis distribution, host star spin rates, and
galactic velocity dispersions have led to estimates for the inspiral
timescale that vary from much less to much greater than the
main-sequence evolution time  ({\it e.g.},
\citealt{jackson_observational_2009}, \citealt{teitler_why_2014},
\citealt{penev_empirical_2018}, \citealt{cameron_hierarchical_2018},
\citealt{hamer_schlaufman_2019}).  Direct measurements of tidal
orbital decay through transit and occultation timing could provide an
empirical solution.  For instance, combined transit timing and radial
velocity measurements for WASP-12b have shown that its secular period
decrease of $\approx$30 milliseconds per year is almost certainly due
to tidal orbital decay
\citep{maciejewski_departure_2016,patra_2017,yee_orbit_2020}.

This study highlights a point that, though obvious, has perhaps not
yet received due attention.  The point is that observational programs
aimed at identifying orbital decay in hot Jupiters through transit
timing will be crippled without concurrent long-term radial velocity
monitoring.  The reason is that line-of-sight accelerations due to
outer companions \citep[{\it e.g.},][]{agol_detecting_2005} and tidal
orbital decay both initially manifest identically in transit times, as
a non-zero period derivative.  Massive outer companions to hot
Jupiters are the norm; \citet{bryan_statistics_2016} calculated an
occurrence rate of $60.9^{+5.2}_{-5.6}\%$ for outer companions to hot
Jupiters with masses from 1-20$M_{\rm Jup}$ and semi-major axes from
5-100$\,$AU.  Therefore it could very well be that more hot Jupiters
will show shrinking orbital periods due to long term accelerations
than due to tidal orbital decay.

The main focus of this study is the hot Jupiter WASP-4b, which has an
orbital period that appears to be decreasing by about 10 milliseconds
per year.  We discovered the timing variations using data from TESS
\citep{ricker_transiting_2015} and a decade of ground-based
observations \citep[][hereafter
\citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}.  Thereafter,
\citet{southworth_transit_2019} reported 22 new transit times for the
system, and found an updated decay rate of $\dot{P} = -9.2 \pm 1.1$
milliseconds per year. The \citeauthor{southworth_transit_2019} decay
rate was $\approx$3$\sigma$ less rapid than that found by
\citetalias{bouma_wasp4b_2019}, but the conclusions of the studies
were otherwise similar.  A separate study by \citet{baluev_2019}
reported additional archival transit light curves of WASP-4b, most
notably from TRAPPIST, and also from select other observers.
\citet{baluev_2019} pointed out that when using lower-precision
subsets of the available transit data, the case for a decreasing
period worsened.  We agree with this statement.

@@

In the following, \S~\ref{sec:observations} collects the available
transit data and presents the new radial velocity and speckle imaging
observations.  \S~\ref{sec:analysis} analyzes the data, and finds that
they yield a picture in which the WASP-4 system is accelerating
towards our line-of-sight, likely due to the pull of a brown or
M-dwarf companion.  \S~\ref{sec:discussion} places this result in the
broader context of orbital decay searches, and points out that
line-of-sight accelerations, {\it i.e.}, ``false positive orbital
decay signals'', are relatively common in the hot Jupiter population.
\S~\ref{sec:conclusions} offers concluding remarks.

@@

% TODO: Details of reduction from Howard or Knutson
After identifying the period decrease in
\citetalias{bouma_wasp4b_2019}, we acquired four additional radial
velocity measurements with the Keck High Resolution Echelle
Spectrometer (HIRES, \citealt{vogt_hires_1994}).  Our observations
were acquired under the purview of the California Planet Survey
\citep{howard_cps_2010}.  The spectra were reduced using {\bf software
X}.  Previously, the HIRES data-points spanned 2010 to 2013
\citep{knutson_friends_2014}.  Our new measurements triple the HIRES
observing baseline to nine years.

@@

% TODO: might Howell suggest a paragraph for how the contrast ratios
% were derived?
We observed WASP-4 twice, on the night of September 11-12 (8 sets of
1000 frames each) and also on the night of September 28-29 (7 sets of
1000 frames each).  {\bf We reconstructed the images following...}
{\bf We reduced the reconstructed speckle images to contrast curves by
injecting and recovering point-sources...} The contrast curves showed
that the second night, which had better seeing, also produces the more
constraining result.  The 832$\,$nm results were also typically the
most constraining, since faint companions are typically redder.  We
therefore opt for the 832$\,$nm September 28-29 results for the
remaining analysis.

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.5\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
  The vertical axis shows the observed transit times minus the
  calculated times assuming a constant orbital period.  More opaque
  points correspond to more precise data.  The quadratic ephemeris and
  its $\pm1\sigma$ statistical uncertainties are shown in blue.
  Neither model fully explains the scatter in transit residuals,
  perhaps due to underestimated uncertainties
  (see~\ref{sec:transit_analysis}).  The binned TESS point (yellow
  star) is the weighted average of 18 TESS transits and is binned for
  display purposes only.  Its uncertainty is slightly larger than the
  marker size.  The models were fitted to all of the individual
  transit times.
  \label{fig:times}
	}
\end{figure}

We considered two models for the observed transit times.  The first
model assumes a constant orbital period $P$ on a circular orbit:
\begin{align}
  t_{\rm tra}(E) &= t_0 + PE,
\end{align}
where $E$ is the epoch number and $t_0$ is a reference epoch.  The
second model assumes that the period changes at a steady rate:
\begin{align}
  t_{\rm tra}(E) &=
    t_0 + PE +
    \frac{1}{2} \frac{{\rm d}P}{{\rm d}E} E^2.
  \label{eq:quadratic_model}
\end{align}
The free parameters are the reference epoch $t_0$, the period at the
reference epoch $P$, and the period derivative, ${\rm d}P/{\rm d}t =
(1/P) {\rm d}P/{\rm d}E$.  We defined the epoch numbers such that
$E=0$ is near the weighted average of the observed times.  This helps
to reduce the covariance between $t_0$ and $P$.  A third possible
model that we did not consider for reasons that will become apparent
is a precessing, eccentric orbit \citep[{\it
e.g.},][]{gimenez_revision_1995,patra_2017}.

@@

From the reduced $\chi^2$ values, we can surmise that neither model
entirely describes the transit data that we have collected---there
must be some additional source of signal or noise. In
\citet{bouma_wasp4b_2019}, we found that the quadratic model for the
(sparser) transit data gave $\chi^2_{\rm red}=1.0$.  The worsened
$\chi^2_{\rm red}$ could reflect underestimated statistical
uncertainties in any of our transit measurements.  It could also
reflect systematic errors in the time-systems in which the transit
measurements were recorded, though we have taken every caution against
this latter possibility.

One approach to analyzing such data would be to introduce a fudge
factor to inflate the transit measurement uncertainties, and lower the
reduced $\chi^2$.  We do not think that such an approach is warranted,
since it would not change the result that the quadratic model is
strongly preferred.

The best-fit period derivative in the quadratic model is
\begin{equation}
\dot{P}
  = - (2.74 \pm 0.28)\times 10^{-10}
  = - 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}.
  \label{eq:dP_dt_obs}
\end{equation}
This agrees to within $1\sigma$ of the value reported by
\citet{southworth_transit_2019} ($\dot{P} = - 9.2 \pm 1.1~{\rm
ms}\,{\rm yr}^{-1}$).  It is $\approx$3$\sigma$ larger than the rate
of period decrease we reported earlier \citep[$- 12.6 \pm 1.2~{\rm
ms}\,{\rm yr}^{-1}$;][]{bouma_wasp4b_2019}, presumably because of the
new data from \citeauthor{southworth_transit_2019} and
\citeauthor{baluev_2019}  The other best-fit transit timing model
parameters are reported in Table~3.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.5\textwidth]{f3.pdf}
	\end{center}
	\vspace{-0.7cm}
    \caption{
      {\bf Zorro contrast curve derived from point-source
      injection-recovery experiments.} Sources below the curve would
      have been detected.  The inset shows the speckle image
      reconstructed from 1000 40 millisecond frames in an 832$\,$nm
      bandpass, and acquired on September 28, 2019.  The image scale
      is $2.46''\times2.46''$.
    }
    \label{fig:zorro}
\end{figure}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.78\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Masses and semi-major axes of companions that meet
    requirements of both the radial velocity and the speckle imaging.}
    The likelihood inferred from radial velocities is shown in
    grayscale, and the region excluded from the speckle imaging is
    indicated (blue line).  The expected mass to semimajor-axis
    degeneracy is shown with a black line.
	\label{fig:mass_sma}
  \vspace{-0cm}
	}
\end{figure*}

@@

We assigned each companion a mass and semimajor axis from log-uniform
distributions within the grid cell. We drew the inclination from a
uniform distribution in $\cos i$.  For companion masses less than
$10\,M_{\rm Jup}$, we drew the eccentricity from
\citet{kipping_beta_2013}'s long-period exoplanet Beta distribution
($a=1.12$, $b=3.09$).  If the companion mass exceeded $10\,M_{\rm
Jup}$, we drew the eccentricity distribution from the power-law $p_e
\propto e^\eta$ reported by \citet{moe_mind_2017} in their
Equation~17.  We emphasize that the long-period exoplanet and
long-period binary eccentricity distributions are quite different: the
exoplanet distribution is ``bottom-heavy'', with eccentricities
preferentially close to zero. The binary star distribution is
``top-heavy'', with eccentricities closer to one.  We chose a mass
cutoff of $10\,M_{\rm Jup}$ to separate the two regimes based on the
bound observed by \citet{schlaufman_evidence_2018} between giant
planets and brown dwarfs, though this value is also close to the
$13\,M{\rm Jup}$ required for a solar-metallicity body to
significantly burn deuterium \citep{burrows_nongray_1997}.

@@

\subsection{Implications for WASP-4}
Previous potential explanations for WASP-4b's decreasing orbital
period included tidal orbital decay, orbital precession, and
light-travel time effects \citep{bouma_wasp4b_2019}.  Our new radial
velocity measurements strongly indicate that the least exotic
option---light-travel time effects---is also the most likely.
Transits show the obital period decreasing by $- 8.64 \pm 0.89~{\rm
ms}\,{\rm yr}^{-1}$; the line-of-sight acceleration observed in radial
velocities would predict a period decrease of $-5.94 \pm 0.39~{\rm
ms}\,{\rm yr}^{-1}$.  Though the quantitative agreement is not
perfect, Occam's razor would suggest that the line of sight
acceleration is probably a sufficient explanation for the apparent
decrease of WASP-4b's orbital period.

The corresponding requirements for the companion causing the
acceleration are that it is likely either a brown-dwarf or low mass
star, orbiting between $10$-$100\,{\rm AU}$ from the host star
(Figure~\ref{fig:mass_sma}).  Given such a mass, this companion could
at one time have influenced the orbital evolution of the inner giant.
The fact that most hot Jupiters have similar massive outer companions
\citep{knutson_friends_2014,bryan_statistics_2016} is circumstantial
evidence for certain high-eccentricity formation channels (see
\citealt{dawson_johnson_2018}).  Further radial velocity monitoring
should eventually reveal the orbital parameters and minimum mass of
WASP-4's companion.

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
  {\bf Predicted hot Jupiter period changes from linear radial
  velocity trends.} Including WASP-4b, 16 of 51 hot Jupiters from
  \citet{knutson_friends_2014} have shown long-term radial velocity
  trends.  The HAT-P-11 signal is plotted, though its radial velocity
  signal may be caused by stellar activity.  Three hot Jupiters are
  not plotted because their radial velocity curves better described as
  quadratic trends in time: HAT-P-17, WASP-8, and WASP-34.  Objects
  are ordered in the $y$ dimension by the absolute value of d$P$/d$t$.
	\label{fig:pdot_pop}
  \vspace{-0.3cm}
	}
\end{figure}

@@

Most hot Jupiters have outer companions with masses larger than
Jupiter beyond 5$\,$AU
\citep{knutson_friends_2014,bryan_statistics_2016}. The accelerations
and period changes induced by these outer companions will become an
increasingly large nuisance in the hunt for orbital decay as the
observational baselines get longer.  In particular, the precision with
which the period derivative can be measured from transits scales as
the fourth power of the baseline duration (\S~\ref{sec:fisher}), and
so within a decade many more hot Jupiters should show orbital period
changes due to accelerations induced by their outer companions.  This
effect can be distinguished from tidal orbital decay through long-term
radial velocity monitoring of hot Jupiters.

>>>>

\begin{abstract}
  Space and ground-based transit observations of the hot Jupiter
  WASP-4b have recently shown that its orbital period is decreasing.
  Proposed explanations for the period change include tidal orbital
  decay, orbital precession, and light-travel time effects.  We
  present new radial velocity measurements of WASP-4, acquired with
  Keck-HIRES.  The data show that the system is accelerating towards
  the Earth at $\dot{\gamma} = -0.0422^{+0.0028}_{-0.0027}\ {\rm
  m}\,{\rm s}^{-1}\,{\rm yr}^{-1}$.  The implied period decrease
  probably explains WASP-4's changing orbital period as a light-travel
  time effect.  Combining the radial velocities with new limits from
  speckle imaging, we find that the system's acceleration is likely
  caused by a 10-200$\,M_{\rm Jup}$ companion with semi-major axis
  between 10-100$\,$AU.  The statistics from
  \citet{knutson_friends_2014} imply that about 1 in 6 hot Jupiters
  are expected to show similar period decreases as WASP-4, due to
  acceleration by outer companions.  We find that these period
  decreases will be increasingly measureable in coming years, because
  the precision of period derivative measurements scales as the fourth
  power of the observing baseline.  Continued radial velocity
  monitoring of hot Jupiters is therefore essential to distinguish
  tidal orbital decay from line-of-sight accelerations.
\end{abstract}

@@

Indirect studies of age and angular momentum indicators including the
hot Jupiter semi-major axis distribution, host star spin rates, and
galactic velocity dispersions have led to estimates for the inspiral
timescale that vary from much less to much greater than the
main-sequence evolution time  ({\it e.g.},
\citealt{jackson_observational_2009}, \citealt{teitler_why_2014},
\citealt{penev_empirical_2018}, \citealt{cameron_hierarchical_2018},
\citealt{hamer_schlaufman_2019}).  Direct measurements of tidal
orbital decay through transit and occultation timing could provide an
empirical resolution.  For instance, combined transit timing and
radial velocity measurements for WASP-12b have shown that its secular
period decrease of $\approx$30 milliseconds per year is almost
certainly due to tidal orbital decay
\citep{maciejewski_departure_2016,patra_2017,yee_orbit_2020}.

This study highlights a point that, though obvious, has perhaps not
yet received due attention.  The point is that observational programs
aimed at identifying orbital decay in hot Jupiters through transit
timing will be crippled without concurrent long-term radial velocity
monitoring.  The reason is that line-of-sight accelerations due to
outer companions \citep[{\it e.g.},][]{agol_detecting_2005} and tidal
orbital decay both initially manifest identically in transit times, as
a non-zero period derivative.  Massive outer companions to hot
Jupiters are the norm; \citet{bryan_statistics_2016} calculated an
occurrence rate of $60.9^{+5.2}_{-5.6}\%$ for outer companions to hot
Jupiters with masses from 1-20$\,$$M_{\rm Jup}$ and semi-major axes
from 5-100$\,$AU.  Therefore it could very well be that more hot
Jupiters will show shrinking orbital periods due to long term
accelerations than due to tidal orbital decay.

The main focus of this study is the hot Jupiter WASP-4b, which has an
orbital period that appears to be decreasing by about 10 milliseconds
per year.  We discovered the timing variations by combining data from
TESS \citep{ricker_transiting_2015} with a decade of ground-based
observations \citep[][hereafter
\citetalias{bouma_wasp4b_2019}]{bouma_wasp4b_2019}.  Thereafter,
\citet{southworth_transit_2019} reported 22 new transit times for the
system, and found an updated decay rate of $\dot{P} = -9.2 \pm 1.1$
milliseconds per year. The \citeauthor{southworth_transit_2019} decay
rate was $\approx$3$\sigma$ less rapid than that found by
\citetalias{bouma_wasp4b_2019}, but the conclusions of the studies
were otherwise similar.  A separate study by \citet{baluev_2019}
reported additional archival transit light curves of WASP-4b from
TRAPPIST and select other observers.  \citet{baluev_2019} pointed out
that when using lower-precision subsets of the available transit data,
the case for a decreasing period worsened.  We agree with this
statement.

@@

In the following, Section~\ref{sec:observations} collects the
available transit data and presents the new radial velocity and
speckle imaging observations.  Section~\ref{sec:analysis} analyzes the
data, and finds that they yield a picture in which the WASP-4 system
is accelerating towards our line-of-sight, likely due to the pull of a
brown or M-dwarf companion.  Section~\ref{sec:discussion} places this
result in the broader context of orbital decay searches, and points
out that line-of-sight accelerations, {\it i.e.}, ``false positive
orbital decay signals'', are relatively common in the hot Jupiter
population.  Section~\ref{sec:conclusions} offers concluding remarks.

@@

After identifying the period decrease in
\citetalias{bouma_wasp4b_2019}, we acquired four additional radial
velocity measurements with the Keck High Resolution Echelle
Spectrometer (HIRES; \citealt{vogt_hires_1994}).  Our observations
were acquired under the purview of the California Planet Survey
\citep{howard_cps_2010}.  The spectra were reduced using {\bf software
X}.  Previously, the HIRES data-points spanned 2010 to 2013
\citep{knutson_friends_2014}.  Our new measurements triple the HIRES
observing baseline to nine years.

@@

% TODO: might Howell suggest a paragraph for how the contrast ratios
% were derived?
We observed WASP-4 twice, on the night of September 11-12 and also
on the night of September 28-29. 
{\bf We reconstructed the images following...}
{\bf We reduced the reconstructed speckle images to contrast curves by
injecting and recovering point-sources...}
The contrast curves showed that the second night, which had better
seeing, also produced the more constraining result.  The 832$\,$nm
results were also the most useful, since faint companions are redder than
the host star.  We therefore opted for the 832$\,$nm September 28-29
results for the remaining analysis.

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.5\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf Timing residuals and best-fit models for WASP-4b.}
  The vertical axis shows the observed transit times minus the
  calculated times assuming a constant orbital period.  More opaque
  points correspond to more precise data.  The quadratic ephemeris and
  its $\pm1\sigma$ statistical uncertainties are shown in blue.
  The scatter in the residuals could be due to underestimated
  measurement uncertainties (see Section~\ref{sec:transit_analysis}).
  The binned TESS point (yellow star) is the weighted average of 18
  TESS transits and is binned for display purposes only.  Its
  uncertainty is slightly larger than the marker size.  The models
  were fitted to all of the individual transit times.
  \label{fig:times}
	}
\end{figure}

We considered two models for the observed transit times.  The first
model assumes a constant orbital period $P$ on a circular orbit:
\begin{align}
  t_{\rm tra}(E) &= t_0 + PE,
\end{align}
where $E$ is the integer transit number and $t_0$ is a reference
epoch.  The second model assumes that the period changes at a steady
rate:
\begin{align}
  t_{\rm tra}(E) &=
    t_0 + PE +
    \frac{1}{2} \frac{{\rm d}P}{{\rm d}E} E^2.
  \label{eq:quadratic_model}
\end{align}
The free parameters are the reference epoch $t_0$, the period at the
reference epoch $P$, and the period derivative, ${\rm d}P/{\rm d}t =
(1/P) {\rm d}P/{\rm d}E$.  We defined the epoch numbers such that
$E=0$ is near the weighted average of the observed times.  This helps
to reduce the covariance between $t_0$ and $P$.  A third possible
model that we did not consider for reasons that will become apparent
is a precessing, eccentric orbit \citep[{\it
e.g.},][]{gimenez_revision_1995,patra_2017}.

@@

From the reduced $\chi^2$ values, we can surmise that neither model
entirely describes the transit data that we have collected---there
must be some additional source of signal or noise. In
\citetalias{bouma_wasp4b_2019}, we found that the quadratic model for the
(sparser) transit data gave $\chi^2_{\rm red}=1.0$.  The worsened
$\chi^2_{\rm red}$ could reflect underestimated statistical
uncertainties in any of our transit measurements.  It could also
reflect systematic errors in the time-systems in which the transit
measurements were recorded, though we have taken every caution against
this latter possibility.

One approach to analyzing such data would be to inflate the transit
measurement uncertainties, and lower the reduced $\chi^2$.  We do not
think that such an approach is warranted, since it would not change
the result that the quadratic model is strongly preferred.

The best-fit period derivative in the quadratic model is
\begin{equation}
\dot{P}
  = - (2.74 \pm 0.28)\times 10^{-10}
  = - 8.64 \pm 0.89~{\rm ms}\,{\rm yr}^{-1}.
  \label{eq:dP_dt_obs}
\end{equation}
This agrees to within $1\sigma$ of the value reported by
\citet{southworth_transit_2019} ($\dot{P} = - 9.2 \pm 1.1~{\rm
ms}\,{\rm yr}^{-1}$).  It is $\approx$3$\sigma$ larger than the rate
of period decrease we previously reported \citepalias[$- 12.6 \pm 1.2~{\rm
ms}\,{\rm yr}^{-1}$;][]{bouma_wasp4b_2019}, presumably because of the
new data from \citeauthor{southworth_transit_2019} and
\citeauthor{baluev_2019}  The other best-fit transit timing model
parameters are reported in Table~3.

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.5\textwidth]{f3.pdf}
	\end{center}
	\vspace{-0.7cm}
    \caption{
      {\bf Zorro contrast limits derived from point-source
      injection-recovery experiments.} Sources below the curve would
      have been detected.  The inset shows the speckle image
      reconstructed from 1000 40 millisecond frames in an 832$\,$nm
      bandpass, and acquired on September 28, 2019.  The image scale
      is $2.46''\times2.46''$.
    }
    \label{fig:zorro}
\end{figure}

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.78\textwidth]{f4.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Masses and semi-major axes of companions that meet
    requirements of both the radial velocities and the speckle
    imaging.} The likelihood inferred from radial velocities is shown
    in grayscale, and the region excluded from the speckle imaging is
    indicated.  The expected mass to semimajor-axis degeneracy is
    shown with a black line.
	\label{fig:mass_sma}
  \vspace{-0cm}
	}
\end{figure*}

@@

We assigned each companion a mass and semimajor axis from log-uniform
distributions within the grid cell. We drew the inclination from a
uniform distribution in $\cos i$.  For companion masses less than
$10\,M_{\rm Jup}$, we drew the eccentricity from
\citet{kipping_beta_2013}'s long-period exoplanet Beta distribution
($a=1.12$, $b=3.09$).  If the companion mass exceeded $10\,M_{\rm
Jup}$, we drew the eccentricity distribution from the power-law $p_e
\propto e^\eta$ reported by \citet{moe_mind_2017} in their
Equation~17.  We emphasize that the long-period exoplanet and
long-period binary eccentricity distributions are quite different: the
exoplanet distribution is ``bottom-heavy'', with eccentricities
preferentially close to zero. The binary star distribution is
``top-heavy'', with eccentricities closer to one.  We chose a mass
cutoff of $10\,M_{\rm Jup}$ to separate the two regimes based on the
bound observed by \citet{schlaufman_evidence_2018} between giant
planets and brown dwarfs, though this value is also close to the
$13\,M_{\rm Jup}$ deuterium-burning limit \citep[{\it
e.g.},][]{burrows_nongray_1997}.

@@

\subsection{Implications for WASP-4}
Previous potential explanations for WASP-4b's decreasing orbital
period included tidal orbital decay, orbital precession, and
light-travel time effects \citep{bouma_wasp4b_2019}.  Our new radial
velocity measurements strongly indicate that the least exotic
option---light-travel time effects---is also the most likely.
Transits show the obital period decreasing by $- 8.64 \pm 0.89~{\rm
ms}\,{\rm yr}^{-1}$; the line-of-sight acceleration observed in radial
velocities would predict a period decrease of $-5.94 \pm 0.39~{\rm
ms}\,{\rm yr}^{-1}$.  Though the quantitative agreement is not
perfect, Occam's razor would suggest that the line-of-sight
acceleration is probably a sufficient explanation for the apparent
decrease of WASP-4b's orbital period.

The corresponding requirements for the companion causing the
acceleration are that it is likely either a brown-dwarf or low mass
star, orbiting between $10$-$100\,{\rm AU}$ from the host star
(Figure~\ref{fig:mass_sma}).  Given such a mass, this companion could
at one time have influenced the orbital evolution of the inner giant.
The fact that most hot Jupiters have similar massive outer companions
\citep{knutson_friends_2014,bryan_statistics_2016} is circumstantial
evidence for certain high-eccentricity formation pathways (see
\citealt{dawson_johnson_2018}).  Further radial velocity monitoring
should eventually reveal the orbital parameters and minimum mass of
WASP-4's companion.

@@

\begin{figure}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.48\textwidth]{f5.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
  {\bf Predicted hot Jupiter period changes from linear radial
  velocity trends.} Including WASP-4b, 16 of 51 hot Jupiters from
  \citet{knutson_friends_2014} have shown long-term radial velocity
  trends.  HAT-P-11 signal is plotted, though its signal may be caused
  by stellar activity.  Three hot Jupiters are not plotted because
  their radial velocity curves better described as quadratic trends in
  time: HAT-P-17, WASP-8, and WASP-34.  Objects are ordered in the $y$
  dimension by the absolute value of d$P$/d$t$.
	\label{fig:pdot_pop}
  \vspace{-0.3cm}
	}
\end{figure}

@@

Most hot Jupiters have outer companions with masses larger than
Jupiter beyond 5$\,$AU
\citep{knutson_friends_2014,bryan_statistics_2016}. The accelerations
and period changes induced by these outer companions will become an
increasingly large nuisance in the hunt for orbital decay as the
observational baselines get longer.  In particular, the precision with
which the period derivative can be measured from transits scales as
the fourth power of the baseline duration (Section~\ref{sec:fisher}), and
so within a decade many more hot Jupiters should show orbital period
changes due to accelerations induced by their outer companions.  
To distinguish this effect from tidal decay, further long-term radial
velocity measurements of hot Jupiters are required.
